---
title: "Bayesian Quantile Tobit regression in Stan"
author: "Miao Cai"
date: "12/23/2018"
output:
  html_document:
    theme: yeti
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
    code_folding: show
  pdf_document:
    number_sections: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```

You need the following packages to run this Rmarkdown file:

- `tidyverse`
- `rstan`
- `brms`
- `shinystan`

# Overview

$$TE = \mathbf{X^{\prime}\beta} + \epsilon \\
\epsilon \sim N(0, \sigma^2)$$

The outcome variable is Technical efficiency (TE). It is assumed to be right-censored at 1. I didn't assume it to be interval-censored since the minimum TE value in your data is more than 0, so no data is left-censored although it could be left-censored theoretically.

The explanatory variables are:

1. bed, 
2. occ, 
3. alos, 
4. dnr, 
5. bnr, 
6. oir, 
7. pop, 
8. urban, 
9. HHI_beds, 
10. gdp

```{r DATA}
library(brms)
library(dplyr)
library(shinystan)

load("data/dat.Rdata")

dat$censoring = 0
dat$censoring[dat$TE_Q == 1] = 1

dat1 = dat %>% 
  select(TE_Q, censoring, occ, bed, alos, dnr, bnr, oir, mc, pop, urban, gdp, bed1000, HHI_beds, year, id) %>% 
  mutate_all(as.numeric)

isFALSE <- function (x) identical(FALSE, x)
```

# Bayesian Tobit regression

## Fixed-effects Bayesian Tobit regression

```{r tobit, eval = FALSE}
fit1 = brm(
  TE|cens(censoring) ~ bed + occ + alos + dnr + bnr + oir + pop + urban + HHI_beds + gdp, 
  family = gaussian, data = dat1, 
  inits = "0", seed = 666, cores = 4,
  chains = 4, iter = 5000, warmup = 2000)

summary(fit1)
save(fit1, file = "fit/fit1.Rdata")
# You can get a matrix of fixef effects coefficients by: 
# fixef(fit1)
readr::write_csv(data.frame(fixef(fit1)), "fit/Fixed-effects Bayesian Tobit regression.csv")

# shinystan::launch_shinystan(fit1)
```

## Bayesian Tobit regression with hospital and year random effects

```{r tobitrandom, eval = FALSE}
fit2 = brm(
  TE|cens(censoring) ~ bed + occ + alos + dnr + bnr + oir + pop + urban + HHI_beds + gdp + (1|id) + (1|year), 
  family = gaussian, data = dat1, 
  inits = "0", seed = 666, cores = 4,
  chains = 4, iter = 5000, warmup = 2000
)

summary(fit2)
save(fit2, file = "fit/fit2.Rdata")
readr::write_csv(data.frame(fixef(fit2)), "fit/Random-effects Bayesian Tobit .csv")
```




A few notes:

- If you want to study the underlying `Stan` codes, you can access them by `stancode(fit2)`.
- If you want to access the priors for this model, you can get them by `get_prior(TE|cens(censoring) ~ bed + occ + alos + dnr + bnr + oir + pop + urban + HHI_beds + gdp + (1|id) + (1|year), family = gaussian, 
  data = dat1)`
- You can use `shinystan::launch_shinystan(fit2)` to explore the posterior  estimates.


# Bayesian quantile Tobit regression

```{r quantiletobit, eval = FALSE}
for (i in 1:9) {
  qfit0 = brm(
  bf(TE_Q|cens(censoring) ~ occ + bed + alos + dnr + bnr + oir + mc + pop + urban + gdp + bed1000 + HHI_beds + (1|id) + (1|year), quantile = 0.1*i), 
  family = asym_laplace(), data = dat1, 
  inits = "0", seed = 666, control = list(max_treedepth = 15), 
  chains = 4, iter = 5000, warmup = 2000, cores = 4
)
  summary(qfit0)
  save(qfit0, file = paste0("fit/fit0", i,".Rdata"))
  readr::write_csv(data.frame(fixef(qfit0)), paste0("fit/fit0", i,".csv"))
}

# You can get a matrix of fixef effects coefficients by: 
# fixef(qfit0)

# shinystan::launch_shinystan(qfit0)
```

## Plotting the quantile Tobit regression results

```{r plot, fig.width=12, fig.height=12/1.618}
require(ggplot2)
require(gridExtra)

#coefficients plot
#read data
for (i in 1:9){
   name <- paste0("temp", i)
   assign(name,read.csv(paste0("fit/fit0", i, ".csv"), header=TRUE))
}

#csvfiles = paste0("fit/", list.files(path = "fit/", pattern = "fit0[0-9].csv"))
#temp = purrr::map(csvfiles, data.table::fread)

#extra data
data_extra <- function(i) {
vars <- c("interecept", "occ", "bed", "alos", "dnr", "bnr", "oir", "mc", "pop", "urban", "gdp", "bed1000", "HHI_beds")
data <- rbind(temp1[i,], temp2[i,], temp3[i,], temp4[i,], temp5[i,], temp6[i,], temp7[i,], temp8[i,], temp9[i,])
data$parname <- vars[i]
data$Quantile <- seq(0.1, 0.9, 0.1)
  return(data)
}
data_extra(1)
#plot function
plot.bqtfit <- function(data) {
  p <- ggplot(data) +
  geom_ribbon(aes(x = Quantile, ymin = Q2.5, ymax=Q97.5), fill="lightblue", alpha = 0.4) + 
  geom_line(aes(x=Quantile, y = Estimate), col= "#0099cc") +
  geom_hline(yintercept = 0, linetype = 2, col="blue") + 
  scale_x_continuous(breaks=seq(0.1, 1, 0.1)) +
  xlab("Quantile") + 
  ylab( paste(toupper(data$parname[1]))) +
  theme_classic() +
  theme(text = element_text(family = "Songti SC", size=6))
  return(p)
}

#plot
#tiff(filename = "Miao/fit/coefplot-bqtfit.tiff", width =10, height = 16, units = "cm", pointsize = 10, bg = "white", res=300)

for (i in 2:13) {
  name <- paste0("p", i)
  assign(name, plot.bqtfit(data_extra(i)))
} 
  grid.arrange(p2, p3, p4, p5, p6, p7, p8, p9, p10, p11, p12, p13,ncol = 2)
```

